### 2023-04-18 thesis/sections/3-exp-setup.tex
basic structure

\section{Experimental Setup}
\label{sec:exp-setup}

\cmtDK[inline]{15 pages}
\cmtDK[inline]{
    - what is ambiguity? \\
    - control over dataset \\
    - target object/target group/non-target group: \\
    ---- 1. how many objects? \\
    ---- 2. how many shared attributes? \\
    ---- 3. which shared attributes? \\
    ---- 4. level of ambiguity \\
    - as close to CLEVR as possible \\
    ---- image size \\
    ---- number (10.000) \\
    ---- between 3 to 10 objects ??? \\
    - unambigous dataset: \\
    ---- target group = non-target group \\
    ---- all attributes shared, except for the color \\
    ---- \ \  --> describable with one 'word'
}