### 2024-02-27 thesis/sections/6-grounding-language-games.tex
thesis

\section{Grounding referring expressions in language games}
\label{sec:language-games}
The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
However, every task is solved by two neural networks that communicate through a discrete bottleneck.
% SD: But there is an importnat difference that now we have two neural netweorks, the sender and the receiver.
% DK: TODO

The language games in this research have an asymmetric setup.
One agent, the sender is shown some information and needs to generate a message.
This message is received by the second agent, the receiver.
The receiver needs to parse this message and combine it with the same information that the sender was presented with.
The receiver then makes a prediction, which is compared to the ground truth.
The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
All weights of the agents are adapted in the same way.

The vocabulary that the sender can draw from to produce a message is made up of initially arbitrary symbols.
The meaning of these symbols is created as soon as the sender uses them in one of the message, and the receiver is able to use it to solve the task successfully.
Over time, specific meanings are reinforced and a language emerges.
Hereby, the production of the messages is constrained in two ways.
First, the size of the vocabulary is varied.
A smaller size forces the agents to compress their information more which is a prerequisite for the emergence of a language.
With more different words, they can express more information, but it is harder to learn.
Different sizes are compared to analyze its effect on the emerged language.
Secondly, the maximum possible length of a message is constrained.
Again, a shorter message requires more compression, while the longer message allows for more information exchange.
% SD: But possibly you would describe the language game setup already before on the background or method chapter. I would move all text there and make the language games description much more detailed, also giving examples. What is happening in terms of interaction? How is this modelled with the neural networks? It is important to emphasise this difference.
% DK: (in both beginning and here, done)

As the experiments before, three different tasks are set up.
For the \emph{object identification} task, the agents play discrimination games similar to the setup in \citep{Lazaridou2017}.
For the \emph{RE generation} task, the receiver needs to generate a human language referring expression based on the emerged language referring expression by the sender.
Since, it is possible that the agents simply learn to align these two languages, the agents are tasked to point towards the center coordinates of the target object in the final task, the \emph{reference resolution} task.
For all three setups, the sender uses the same setup and only the receiver is switched.
By this, differences in the produced referring expression in the message by the sender are only dependent on the receiver's ability to solve the different tasks.
In particular, the impact of the task on the emerged language can be analyzed in more detail.

As discussed in Section \ref{sec:background:lg:setup}, language games are evaluated on the success of communication and the emergence of a language.
The testing phase is not necessarily used to test the generalization capabilities of the agents, but rather to extract a sample of the emerged language.
It is therefore not as important that the samples in the testing phase are unseen samples.
Hence, for both training and testing a certain number of samples are randomly drawn from the complete dataset.
In these experiments, the dataset consists of 10.000 scenes in total.
To train the agents, 128.000 games are played (using 128.000 randomly drawn samples from the dataset) with a batch size of 32 samples.
To retrieve the samples of the emerged language, after the agents are trained, they are frozen and play 2.048 games with randomly drawn samples from the dataset.
This yields 2.048 interactions that will be used to analyze the emerged language.

Across all experiments, the same restriction on the vocabulary and the produced messages are compared.
First, five different vocabulary sizes $|V|$ are tested: 2, 10, 16, 50 and 100 symbols.
Notice that these vocabularies include the \emph{<eos>} token which signals the end of the message.
A vocabulary of two symbols therefore consists effectively of only one token that can be learned by the agents.
The center of attention lies on the vocabularies with 16 and 50 symbols.
The vocabulary with 16 symbols corresponds the 13 attributes the objects can have and align with human language referring expressions.
However, a smaller vocabulary size increases the pressure to converge on meaning, which is my also 10 and 2 symbols are tested.
On the other side, the agents might instead of a compositional language learn to assign one symbol with every combination of attributes.
This results in $2 * 3 * 8 = 42$ unique combinations, which corresponds to the vocabulary with 50 symbols.
Furthermore, a vocabulary of 100 symbols is used to test the effect of a much bigger bottleneck.

Secondly, also different maximum message lengths are compared: 1, 2, 3, 4 and 6 symbols.
Notice that this applies only to the maximum possible length; the agents can choose to use shorter messages.
Again, the corresponding length to human referring expressions in English is 3 for the three attributes.
However, the agents can be pressured to condense the information more precisely with shorter messages or given less restrictions with longer messages.
Vocabulary size and message length play hereby an opposing role in the pressure on condensing information.
% If the symbols were similarly used, messages would have lengths between one and three symbols.
% Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
% Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
% Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
% DK: TODO

All agents are trained using the \emph{Adam} optimizer \citep{Kingma2015}.
LSTMs are used to produce and parse the message between the agents.
Gumbel-Softmax relaxation is applied to train the agents with those discrete messages.
Hereby, a temperature of $\tau=1$ is used which showed the best results in preliminary experiments.
All setups are additionally compared to a baseline.
In the baseline the sender is "blinded", meaning that the sender is not shown any input but produces random messages.
Effectively, the receiver needs to solve the task on their own, by exploiting underlying structures in the shown data and the target data.
If a non-baseline model doesn't achieve a better performance than the baseline, the sender didn't communicate any meaningful messages and no language emerged.